---
title: "Gov 50 Final Project"
author: "Addie Rose Bullock"
description: "How did the Quality of Democracy change for NATO countries after the 9/11 Terrorist Attacks on the United States"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction:

In my final project, I decided to look into the quality of democracy in countries when democracy is challenged or attacked. I will use the attacks on the United States on September 11th, 2001 as the event, and will look at the United States and the 11 other original (NATO) North Atlantic Treaty Organization members (Belgium, Canada, Denmark, France, Iceland, Italy, Luxembourg, Netherlands, Norway, Portugal, United Kingdom, and United States). Since NATO was formed to promote freedom in 1949, I felt that these 12 countries accurately support and defend democracy. I predict that after 9/11, the levels and views of democracy in the 12 original NATO countries will increase, due to the level of threat presented. I think this is interesting because of a large attack on the United States can cause ripple of effects throughout the world.


```{r}
library(readr)
library(tidyverse)
library(broom)
library(readxl)
db_data <- read_xlsx('DB_data_2020_all_data.xlsx')
db_data_sd <- read_xlsx('DB_data_2020_all_data_standardised.xlsx')
```


```{r}
db_data_nato <- db_data_sd  |>
  filter(Country  %in%  c("Belgium","Canada","Denmark","France","Iceland","Italy","Luxembourg","Netherlands","Norway","Portugal","United Kingdom","United States")) |>
  filter(Year >=  1995 & Year <= 2005)|>
  select(Country, Year, Countryyear, FREEDOM, CONTROL, EQUALITY, DQ)  |>
  mutate(is_US = ifelse(Country == "United States", "United States", "Other"),
         after_911 = ifelse(Year > 2001, "Post 9/11", "Before 9/11"))
```

## Data:

```{r}
db_data_nato |>
  ggplot(aes(x = DQ)) +
  geom_histogram(fill = 'red') +
  labs(x = 'Quality of Democracy',
       y = 'Count',
       title =  'Quality of Democracy in NATO Countries from 1995-2005')

summary(db_data_nato$DQ)


```
This data is from the democracy barometer, which determines the quality of democracy by three sections: freedom, control, and equality. I explored the data with a before-and-after design with the key date being Septermber 11th,  2001. For each individual index, it requires expert coding to determine each level to calculate the quality of democracy. Freedom is defined as the individual's liberties, rule of law, and public sphere. Control is defined as the competition, market, mutual constraints, and governmental capability. Equality is defined as the transparency, participation, and representation of the constituents. The quality of democracy is an average of these things, but is on an unit-less scale. The codebook states that researchers can create their own scale using the raw measurement, but we can use the raw data to compare between countries and over time. My dependent variable will be the Quality of Democracy (DQ) from 1995-2005, while the independent variables are year, before and after 9/11, and whether or not the country is the United States. I will measure this by looking at the change over 10 years. 

Democracy Barometer Website:
https://democracybarometer.org/data-and-documentation/

## Results:

```{r}

db_data_nato |>
  ggplot(aes(x = as.numeric(Year), y = DQ, group = Country))  +
  geom_line(aes(color = Country))  +
  geom_vline(mapping = aes(xintercept = 2001.696)) +
  scale_color_manual(values = c("Belgium" = 'green',"Canada" = 'orange',"Denmark" = 'yellow',"France" =  'red',"Iceland" = 'light blue',"Italy" = 'pink',"Luxembourg" = 'purple',"Netherlands" = 'dark blue',"Norway" = 'hot pink',"Portugal" = 'lime green',"United Kingdom" = 'dark green',"United States" = 'black')) +
  scale_x_continuous(breaks = c(1995:2005), labels = c(1995:2005))  +
  labs(x = 'Year',
       y = 'Quality of Democracy',
       title = 'Change in Quality of Democracy in NATO Countries from 1995 to 2005')

```

```{r}

lm2 <- lm(data  = db_data_nato,
          formula = DQ ~ as.numeric(Year)*after_911*is_US)

d <- coef(summary(lm2))
rownames(d) <- c("Intercept", "Year", "After 9/11",
                 "United States", "Year:After 9/11",
                 "Year:United States", "After 9/11:United States",
                 "Year:After 9/11:United States")

d|>
  knitr::kable(digits = 3)

db_data_nato$preds <- predict.lm(lm2 , newdata =  db_data_nato) 

db_data_nato |>
  mutate(groups = paste0(as.character(is_US),as.character(after_911))) |>
  ggplot(aes(x  =as.numeric(Year), y= preds)) +
  geom_line(aes(color = groups)) +
  geom_vline(xintercept =2001.696) +
  scale_x_continuous(breaks = c(1995:2005), labels = c(1995:2005))  +
  labs(x = 'Year',
       y = 'Quality of Democracy',
       title = 'Predicted Change in Quality of Democracy\nin US vs. NATO Countries Before and After 9/11')


```
Because we have two binary variables (post-9/11 and United States), we can look at the regression model as four separate equations, each predicting Democracy Quality by Year, with varying intercepts and slopes:

* Before 9/11 & the US: Predicted QD = 36.59235 - 0.016338(Year)
* After 9/11 & the US: Predicted QD = 1.505833 + 0.001147(Year)
* Before 9/11 & not the US: Predicted QD = -11.977289 + 0.008037(Year)
* After 9/11 & not the US: Predicted QD = 21.04051 - 0.008466(Year)

The plot above shows these predicted results, based on the regression. Visually, we see that before 9/11, the Quality of Democracy was declining in the United States and increasing in that other NATO countries. Visually, after 9/11, we see that Quality of Democracy drops drastically for the U.S. and begins to incline again, while the NATO begins to decline. Because I wanted varying slopes and varying intercepts, there is not one main coeffient due to the interaction terms. Which is why I have provided the four equations above. We can see that after 9/11 there is a positive coefficient on year for the U.S., indicating that on average the DQ score increased by 0.01147. We do see that in the United States, the drastic drop in intercept did fall from 36.59235 to 1.505833. None of these results were statistically significant, as the p-values were all far above the alpha level of 0.05. 

I do not believe this to be a causal relationship, as there are many possible confounders and omitted variables. For instance,  there may have been other foreign or domestic policies in the U.S. or  other NATO  countries that were enacted around the same as 9/11, but were unrelated to the terrorist attack. 


## Conclusion:

In this study, I found that there is no statistical significance of changes democracy score over time for either the US or non-US NATO countries, nor are there statistically significant differences in democracy score between the US and NATO countries before or after 9/11. Because none of the coefficeiints are significant for the alpha = 0.05, I cannot reject the null hypothesis that democracy in these 12 countries did not increase. A limitation was making the proper regression that could accurately compare, so I would explore doing a true time series in the future. A limitation of my analysis was not looking into the public's opinion of democracy in these countries, because the reaction of the people is immediate, while the reactions of the government take time.


